FAIR Theory: Applying Open Science Principles to the Construction and Iterative Improvement of Scientific Theories

### 2024-12-19 manuscript.Rmd

author:
  - name: "Caspar J. Van Lissa"
    affiliation: "1"
    corresponding: yes
    address: "Professor Cobbenhagenlaan 125, 5037 DB Tilburg, The Netherlands"
    email: "c.j.vanlissa@tilburguniversity.edu"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Formal Analysis
      - Funding acquisition
      - Methodology
      - Project administration
      - Software
      - Supervision
      - Writing – original draft
      - Writing – review & editing
  - name: "Aaron Peikert"
    affiliation: "2,3"
    role:
      - Formal Analysis
      - Writing – original draft
      - Writing – review & editing
  - name: "Andreas M. Brandmaier"
    affiliation: "2,3,4"
    role:
      - Formal Analysis
      - Writing – original draft
      - Writing – review & editing
  - name: "Felix Schönbrodt"
    affiliation: "5"
    role:
      - Conceptualization
      - Writing – review & editing
affiliation:
  - id            : "1"
    institution   : "Tilburg University, dept. Methodology & Statistics"
  - id            : "2"
    institution   : "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
  - id            : "3"
    institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany"
  - id            : "4"
    institution   : "Department of Psychology, MSB Medical School Berlin, Berlin, Germany"
  - id            : "5"
    institution   : "Other affiliations"
authornote: |
  This is a preprint paper, generated from Git Commit # `r substr(gert::git_commit_id(),1,8)`.

@@

Given these concerns, it is unfortunate that scientific reform initiated by the open science movement has focused primarily on improving deductive methods.
The equally critical inductive processes of theory construction and theory improvement have been largely overlooked.
The present paper restores balance by applying, for the first time,
open science principles to psychological theory.
We apply the FAIR principles to scientific theories,
introducing the concept of *FAIR theory* to 
facilitate transparent scholarly communication and accelerate cumulative knowledge acquisition.

>>>>

author:
  - name: "Caspar J. Van Lissa"
    affiliation: "1"
    corresponding: yes
    address: "Professor Cobbenhagenlaan 125, 5037 DB Tilburg, The Netherlands"
    email: "c.j.vanlissa@tilburguniversity.edu"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Formal Analysis
      - Funding acquisition
      - Methodology
      - Project administration
      - Software
      - Supervision
      - Writing – original draft
      - Writing – review & editing
  - name: "Aaron Peikert"
    affiliation: "2,3"
    role:
      - Formal Analysis
      - Writing – original draft
      - Writing – review & editing
  - name: "Andreas M. Brandmaier"
    affiliation: "2,3,4"
    role:
      - Formal Analysis
      - Writing – original draft
      - Writing – review & editing
  - name: "Felix D. Schönbrodt"
    affiliation: "5"
    role:
      - Conceptualization
      - Writing – review & editing
affiliation:
  - id            : "1"
    institution   : "Tilburg University, dept. Methodology & Statistics"
  - id            : "2"
    institution   : "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
  - id            : "3"
    institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany"
  - id            : "4"
    institution   : "Department of Psychology, MSB Medical School Berlin, Berlin, Germany"
  - id            : "5"
    institution   : "Ludwig-Maximilians-Universität München, Germany"
authornote: |
  This is a preprint paper, generated from Git Commit # `r substr(gert::git_commit_id(),1,8)`.

@@

Given these concerns, it is an imbalance that scientific reform initiated by the open science movement has focused primarily on improving deductive methods.
The equally critical inductive processes of theory construction and theory improvement have been largely overlooked.
The present paper restores balance by applying, for the first time,
open science principles to psychological theory.
We apply the FAIR principles to scientific theories,
introducing the concept of *FAIR theory* to 
facilitate transparent scholarly communication and accelerate cumulative knowledge acquisition.
